---
title: "Project-2"
output:
  word_document: default
  pdf_document:
    number_sections: yes
    latex_engine: xelatex
fig_caption: yes
---


```{r setup, include=FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE)
```

```{r,echo = TRUE,message=FALSE}
library(dplyr)
library(moderndive)
library(ISLR)
library(skimr)
library(plotly)
library(tidyr)
library(jtools)
library(kableExtra)
#install.packages("corrplot")
library(corrplot)
library(ggplot2)
library(janitor)
```

# Introduction
In this analysis, we are going to analysis which variables have influence on the member numbers of households in the region of **Northern Mindanao** in Philippines.

Since the response variable is a count data, not continuous any more. It is suitable to fit a GLM model, particularly a Poisson model.

# Exploratory analysis

```{r,}

# Dataset prepare
data <- read.csv("dataset3.csv")

# Ignore the region and rename
household <- data[,-2]
colnames(household) <- c("Income", "Food.exp", "Head.sex", "Head.age", "Type", "Members.no", "Area", 
                         "House.age", "Bedrooms.no", "Elec.")

# Check is there any NA in data
cat("\n Check is there any NA in data \n")
apply(household, 2, function(x) any(is.na(x))) # No NAs in data

# Move our response variable to the first
household <- household %>%
  relocate("Members.no") # Use dplyr::relocate to move to the front
head(household) %>%   
  kable(digits = 2) %>%
  kable_styling(latex_options = 'HOLD_position')

```

```{r, fig.width = 6, fig.height = 4, fig.align = "center", fig.pos = "H"}
# summary of response varibale
household %>% select(Members.no)%>%
  skim()
unique(household$Members.no)

max.no <- max(household$Members.no)
ggplot(household, aes(x = Members.no)) +
  # Set bins as max(Members.no)
  geom_histogram(bins = max.no, fill = "#0038A8", col = "white") +
  labs(main = "Total number of family members in Northern Mindanao")

# We first cut Y into classes
median.no <- quantile(household$Members.no, probs = c(0.5));median.no
# Median.no is 4

# Binary classification
points1 <- c(1, 4, 16)
# Multiclass classification (n = 4)
points2 <- c(1, 4, 8, 12, 16)

household$quantiles <- cut(household$Members.no, breaks = points1,
                           include.lowest = TRUE, labels = c("1-4", "5-16"))
# household$quantiles <- cut(household$Members.no, breaks = points2,
#                           include.lowest = TRUE, labels = c("1-4","5-8","9-12","13-16"))

household %>% 
  group_by(quantiles) %>%
  summarize(n=n())

```


## Check our categorical variables with Y
### Check Head.sex vs. Member.no

```{r,}
# Code from lab9
# use the janitor package to summarise this data in a table format:
household %>% 
  tabyl(Head.sex, quantiles) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() # To show original counts

p1 <- ggplot(household, aes(x= as.factor(quantiles),  y = ..prop.., group=Head.sex, fill=Head.sex)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion",x = "", main = "Barplot of Members.no by Head.sex")
p1
```

```{r,}
# Check Type
household %>% 
  tabyl(Type, quantiles) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() # To show original counts

p2 <- ggplot(household, aes(x= as.factor(quantiles),  y = ..prop.., group=Type, fill=Type)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion",x = "", main = "Barplot of Members.no by type")
p2
```

```{r,}
# Check Electricity
household %>% 
  tabyl(Elec., quantiles) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() # To show original counts

p3 <- ggplot(household, aes(x= as.factor(quantiles),  y = ..prop.., 
                            group=as.factor(Elec.), fill=as.factor(Elec.))) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion",x = "", main = "Barplot of Members.no by electricity")
p3
```


Short conclusion:

- For *Head.sex*, a large proportion in *1-4*-member family are female (67.8% vs 32.2%), while in *5-16* more families are with male heads (53.7% vs 46.3%)
- For *Type*, ...
- For *Electricity*, ...
(... need to be done)


## Check the relationships between numerical variables

### We first check correlations betwwen Xs
```{r, eval = TRUE, out.width = '75%', fig.align = "center", fig.cap = "\\label{fig:cor} Correlations of all variables.", fig.pos = "H"}
library(GGally)

ggpairs(household[,c("Members.no", "Income", "Food.exp", "Head.age", "Area", "House.age", "Bedrooms.no")]) 
```
We notice a high cor between *Income* and *Food.exp*. (0.792)


### Check Y vs. Xs one by one
```{r, }
## Why we use log-transform

# Plotting with continuous variables
# Income, Food.exp, head.age
# ------------------ Income------------------------- #
p1 <- ggplot(data = household, aes(x = as.factor(quantiles), y = Income, fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Income")+ 
    theme(legend.position = "none", title = element_text(size = 16)) +
    labs(title = "Members number by Income")

p1.log <-ggplot(data = household, aes(x = as.factor(quantiles), y = log(Income), fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Log Income")+ 
    theme(legend.position = "none", title = element_text(size = 16)) +
    labs(title = "Members number by log Income")
p1; p1.log
```

We applied log-transform for *Income*, to stable the variability.
Similarly, we did to *Food.exp*, *Area* and *House.age*


```{r, fig.width = 9, fig.height = 12, fig.align = "center", fig.pos = "H"}}
library(gridExtra)

# Plotting with numerical variables
# Income, Food.exp, head.age
# ------------------ Income------------------------- #
p1.log <-ggplot(data = household, aes(x = as.factor(quantiles), y = log(Income), fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Log Income")+ 
    theme(legend.position = "none") +
    labs(title = "Members number by log income")

# ------------------ Food.exp ------------------------- #
p2.log <- ggplot(data = household, aes(x = as.factor(quantiles), y = log(Food.exp), fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Log food.exp")+ 
    theme(legend.position = "none") +
    labs(title = "Members number by log food.exp")


# ------------------ Head.age ------------------------- #
p3 <- ggplot(data = household, aes(x = as.factor(quantiles), y = Head.age, fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Head.age")+ 
    # Add a mean age line
    geom_hline(yintercept = mean(household$Head.age), 
               linetype = "dashed") +
    theme(legend.position = "none") +
    labs(title = "Members number by head age")

grid.arrange(p1.log, p2.log, p3,  ncol = 3)
```

```{r, fig.width = 12, fig.height = 12, fig.align = "center", fig.pos = "H"}}
library(gridExtra)

# Plotting with continuous variables
# Area, House.age, Bedrooms.no
# ------------------ Area ------------------------- #
p1.log <-ggplot(data = household, aes(x = as.factor(quantiles), y = log(Area), fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Log Area")+ 
    theme(legend.position = "none") +
    labs(title = "Members number by log area")
# ------------------ House.age ------------------------- #
p2.log <- ggplot(data = household, aes(x = as.factor(quantiles), y = log(House.age), fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Log house.age")+ 
    theme(legend.position = "none") +
    labs(title = "Members number by log house age")

# ------------------ Bedrooms.no ------------------------- #
p3 <- ggplot(data = household, aes(x = as.factor(quantiles), y = Bedrooms.no, fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Bedrooms.no")+ 
    theme(legend.position = "none") +
    labs(title = "Members number by bedroom number")

grid.arrange(p1.log, p2.log, p3,  ncol = 3)

```

Short conclusion:

- For *income*, ...
- For *food.exp*, ...
- For *head age*, ...
- For *house age*, ...
- For *area* and *bedrooms number*,  there is no significantly difference.
(... need to be done)


# Correlation of all

```{r,  fig.width = 9, fig.height = 6, fig.align = "center", fig.pos = "H"}


household <- mutate(household, 
                    Income.log = log(Income),
                    Food.exp.log = log(Food.exp),
                    House.age.log = log(House.age))

#Corrplot
household %>%
    select(Members.no, Income.log, Food.exp.log, Head.age, House.age.log) %>%
    cor() %>%
    # Code reference:
    # http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram
    corrplot(type="upper",
             addCoef.col = "black", # Add coefficient of correlation
             tl.col="black", tl.srt=45, #Text label color and rotation
             diag=FALSE) # hide correlation coefficient on the principal diagonal

```
# Explain of Nan of log form

There is an abs. with house age ZERO, result in an INF in house.age.log

```{r, fig.width = 9, fig.height = 6, fig.align = "center", fig.pos = "H"}

# We ignore that Inf data
household <- household[!is.infinite(household$House.age.log),]
nrow(household) #1884

# Get the correlation pattern
ggpairs(household[,c("Income.log", "Food.exp.log", "Head.age", "House.age.log")]) 
```
```
Short conclusion:

Among numerical variabls, *Income*, *Food.exp*, *Head.age*, *House.age* may be better.
Moreover, *Income* and*Food.exp* are highly correlated. Pay attention.



# Modelling
Note by Huan: Maybe Poisson model is more appropriate? The meeting can be discussed together~~
The modelling details are all there, confidence intervals and predictions are done, thanks for xuan working hard!

```{r,echo = FALSE}
library("janitor")

data_cut <- household[order(household$Members.no, decreasing = T),]
data_cut <- data_cut %>%
    mutate(Num2 = data$Total.Number.of.Family.members)
data_cut[,"Num2"] <- ifelse(data_cut[,"Num2"] > median(data_cut[,"Num2"]),"High","Low")

data_sex2 <- data_cut %>%
    select(Num2, Household.Head.Sex)

#look
data_sex2 %>% 
  tabyl(Household.Head.Sex, Num2) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

ggplot(data_sex2, aes(x= Num2,  y = ..prop.., group=Household.Head.Sex, fill=Household.Head.Sex)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion")

model_sex2 <- glm(Num2 ~ Household.Head.Sex, data = data_sex2, family = binomial(link = "logit"))
model_sex2 %>%
  summary()
levels(data_sex2$Household.Head.Sex)
confint(model_sex2) %>%
  kable()

mod.ethnic.coef.logodds <- model_sex2 %>%
                            summary() %>%
                            coef()

mod.ethnic.coef.logodds <- model_sex2 %>%
                            summary() %>%
                            coef()

mod.ethnic.coef.logodds <- model_sex2 %>%
                            summary() %>%
                            coef()

plot_model(model_sex2, show.values = TRUE, transform = NULL,
           title = "Log-Odds (Male instructor)", show.p = FALSE)
```

#Cut 5
```{r,echo = FALSE}
data_cut <- read.csv("C:/Users/Seismic/Downloads/dataset3.csv")
data_cut <- data_cut %>%
    mutate(Num5 = data$Total.Number.of.Family.members)
data_cut<-data_cut[order(data_cut$Num5,decreasing = T),]
data_cut$Num5<-cut(data_cut$Num5,breaks=c(1,3,6,10,16),labels=c('low1','low2','high1','high2'))



model_sex3 <-  multinom(Num5 ~ Household.Head.Sex, data = data_cut, probabilities = TRUE, model = TRUE)
summary(model_sex3)
Anova(model_type)
round(odds.ratio(model_type),2)

```

#Poisson model
```{r}

```

